{\small
\section*{Outline and Due Dates}
\red{Also add a list of figures???}
\red{\\ Oliver will probably want to see pictures of scalings and timings. Scaling/processor etc. Include some nice pictures of these things.}
\begin{enumerate}
   \item Background/Motivation \red{Tuesday Feb 19}
   \red{Hardest section but really important. Make sure to explain why the stuff we're working on is interesting, this will lead into talking about other methods.}
   \begin{enumerate}
      \item Other methods like HF and no-core shell model
      \item General outline of the dissertation
   \end{enumerate}
   \item Quantum Monte Carlo \red{Friday Feb 22}
   \begin{enumerate}
      \item VMC
      \item DMC
      \item AFDMC
      \begin{enumerate}
         \item Mixed expectation values, operator breakup
      \end{enumerate}
      \item Hamiltonian - Phenomenological + $\chi$EFT + some results with AFDMC/GFMC
   \end{enumerate}
   \item Trial Wave Function \red{Friday Feb 15}
   \begin{enumerate}
      \item Slater Determinants
      \item Pfaffian
      \red{Explain what it's used for (superfluidity etc.) and why it's more efficient than the SD stuff (need many thousands of SD's to do the same thing).}
      \item Spin-Isospin Dependent Correlations
      \begin{enumerate}
         \item Quadratic Correlations - include results
         \begin{enumerate}
            \item Results
         \end{enumerate}
         \item Exponential correlations \ldots why they aren't working or how to do it, cluster decomposition
         \red{If not working then don't say much about it. Explain that there are large variances that need to be taken care of etc.}
      \end{enumerate}
      \item Alessandro's wave function with $T^2$ fix attempts. \blue{Maybe \ldots maybe don't include}
   \end{enumerate}
   \item Alpha-clustering \red{Friday Mar 1}
   \begin{enumerate}
      \item Theory
      \item Results
   \end{enumerate}
   \item Conclusion \red{Friday Mar 8}
   \begin{enumerate}
      \item Future Work
   \end{enumerate}
\end{enumerate}
\blue{Make sure to include stuff for each member of the committee.}
}
