\begin{abstract}
Quantum Monte Carlo is one of the most accuracte {\it ab initio} methods used to study nuclear physics. The accuracy and efficiency depend heavily on the trial wave function, especially in Auxiliary Field Diffusion Monte Carlo (AFDMC), where a simplified wave function is often used to allow calculations of larger systems. The simple wave functions used with AFDMC contain short range correlations that come from an expansion of the full correlations truncated to linear order. I have extended that expansion to quadratic order in the pair correlations. I have investigated this expansion by keeping the full set of quadratic correlations as well an expansion that keeps only independent pair quadratic correlations. To test these new wave functions I have calculated ground state energies of $^4$He, $^{16}$O, $^{40}$Ca and symmetric nuclear matter at saturation density $\rho=$0.16 fm$^{-3}$ with 28 particles in a periodic box. The ground state energies calculated with both wave functions decrease with respect to the simpler wave function with linear correlations only for all systems except $^4$He for both variational and AFDMC calculations. It was not expected that the ground state energy of $^4$He would decrease due to the simplicity of the alpha particle wave function. These correlations have also been applied to study alpha particle formation in neutron rich matter, with applications to neutron star crusts and neutron rich nuclei. I have been able to show that our method can be used to study small clusters as well as the effect of external nucleons on these clusters.
\end{abstract}
