\begin{acknowledgements}
I have had many academic advisers during this journey, but I will mention only a few here. My undergraduate physics adviser, Steve Turley, who always believed I could easily handle any task he gave me, even if I didn't believe I could. Thomas Leitner, my summer internship adviser at Los Alamos National Laboratory who has advised me as much in science as he has in other aspects of life. My graduate school adviser, Kevin Schmidt, for the example he has set in being an excellent research scientist. From Kevin I have learned how to approach problems on my own and build my understanding from the bottom up.

I would also like to thank those on my committee, Igor Shovkovy, Oliver Beckstein, and Ricardo Alarc\'on for their time, helpful advise toward my research, as well as the advise and encouragement they have given me toward other aspects of my life. I am also indebted to the ASU physics staff, especially Araceli Vizcarra, who always went above and beyond to help with everything.

I am also grateful to my collaborators Stefano Gandolfi and Joe Carlson for helping me develop as a young scientist. I am also thankful for the advise and discussions with many of my friends and group members, many of whom could also be considered collaborators, specifically Lucas Madeira, Rong Chen, Diego Lonardoni, Alessandro Roggero, Lorenzo Andreoli, and Andy Svesko.

I gratefully acknowledge funding support from National Science Foundation Grant No. PHY-1404405 and by a Wally Stoelzel Fellowship from the ASU physics department. This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number ACI-1548562.
\end{acknowledgements}
